
---
author: |
  | William Findlay
title: |
  | RELI2732 Summary Notes
date: \today
bibliography: /home/housedhorse/.bibs/uni.bib
csl: /home/housedhorse/.bibs/ieee.csl
subparagraph: yes
header-includes: |
  ```{=latex}
  \usepackage{float}
  \usepackage{listings}
  \usepackage[hang,bf]{caption}
  \usepackage{framed}
  \usepackage[section]{placeins}

  \usepackage[dvipsnames,table]{xcolor}

  \usepackage{pifont}

  \allowdisplaybreaks

  % fancy headers/footers
  \makeatletter
  \usepackage{fancyhdr}
  \lhead{\@author}
  \chead{}
  \rhead{\@title}
  \lfoot{}
  \cfoot{\thepage}
  \rfoot{}
  \renewcommand{\headrulewidth}{0.4pt}

  \usepackage{amsmath, amsfonts,amssymb, amsthm}
  \usepackage{siunitx}
  \usepackage[boxruled,lined,linesnumbered,titlenumbered]{algorithm2e}

  \usepackage{setspace}
  \usepackage{changepage}
  \usepackage[explicit]{titlesec}
  \usepackage{aliascnt}

  \floatplacement{figure}{!htbp}
  \floatplacement{table}{!htbp}
  \lstset{numbers=left,breaklines=true,frame=single,language=Python,captionpos=t,abovecaptionskip={\abovecaptionskip},
  belowcaptionskip={0.5em},aboveskip=\intextsep,showstringspaces=false,identifierstyle=\color{Blue},
  commentstyle={\color{OliveGreen}},keywordstyle={\bfseries\color{Orange}}}
  \setlength{\captionmargin}{1in}

  \newgeometry{margin=1in}

  \newtheoremstyle{plain}
  {12pt}   % ABOVESPACE
  {12pt}   % BELOWSPACE
  {\itshape}  % BODYFONT
  {0pt}       % INDENT (empty value is the same as 0pt)
  {\bfseries} % HEADFONT
  {.}         % HEADPUNCT
  {5pt plus 1pt minus 1pt} % HEADSPACE
  {}          % CUSTOM-HEAD-SPEC

  \newtheoremstyle{definition}
  {12pt}   % ABOVESPACE
  {12pt}   % BELOWSPACE
  {\normalfont}  % BODYFONT
  {0pt}       % INDENT (empty value is the same as 0pt)
  {\bfseries} % HEADFONT
  {.}         % HEADPUNCT
  {5pt plus 1pt minus 1pt} % HEADSPACE
  {}          % CUSTOM-HEAD-SPEC

  \newtheoremstyle{remark}
  {12pt}   % ABOVESPACE
  {12pt}   % BELOWSPACE
  {\normalfont}  % BODYFONT
  {0pt}       % INDENT (empty value is the same as 0pt)
  {\itshape} % HEADFONT
  {.}         % HEADPUNCT
  {5pt plus 1pt minus 1pt} % HEADSPACE
  {}          % CUSTOM-HEAD-SPEC

  \theoremstyle{plain}

  % define theorem
  \newtheorem{theorem}{Theorem}[section]
  \providecommand*{\theoremautorefname}{Theorem}

  % define lemma
  \newtheorem{lemma}{Lemma}[section]
  \providecommand*{\lemmaautorefname}{Lemma}

  % define claim
  \newtheorem{claim}{Claim}[section]
  \providecommand*{\claimautorefname}{Claim}

  % define corollary
  \newtheorem{corollary}{Corollary}[section]
  \providecommand*{\corollaryautorefname}{Corollary}

  % define proposition
  \newtheorem{proposition}{Proposition}[section]
  \providecommand*{\propositionautorefname}{Proposition}

  % define conjecture
  \newtheorem{conjecture}{Conjecture}[section]
  \providecommand*{\conjectureautorefname}{Conjecture}

  \theoremstyle{remark}

  % define observation
  \newtheorem{observation}{Observation}[section]
  \providecommand*{\observationautorefname}{Observation}

  % define remark
  \newtheorem{remark}{Remark}[section]
  \providecommand*{\remarkautorefname}{Remark}

  \theoremstyle{definition}

  % define example
  \newtheorem{example}{Example}[section]
  \providecommand*{\exampleautorefname}{Example}

  % define definition
  \newtheorem{definition}{Definition}[section]
  \providecommand*{\definitionautorefname}{Definition}

  \newcommand{\blackbox}{\hfill$\blacksquare$}
  \usepackage{tikz}
  \newcommand*\circled[1]{\tikz[baseline=(char.base)]{
              \node[shape=circle,draw,inner sep=2pt] (char) {#1};}}

  \renewcommand{\labelitemi}{\large$\bullet$}
  \renewcommand{\labelitemii}{\ding{226}}
  \renewcommand{\labelitemiii}{\tiny$\blacksquare$}
  \renewcommand{\labelitemiv}{\small$\triangleright$}

  \titleformat{\paragraph} % command to change
  [runin]                  % shape  (runin, etc.)
  {\bfseries}              % format (bfseries, itshape, etc.)
  {}                       % label  (thesection, thesubsection, etc.)
  {0em}                    % separation between label and body
  {#1}        % before the body
  [.]                       % after the body

  \titleformat{\subparagraph} % command to change
  [runin]                  % shape  (runin, etc.)
  {\itshape}              % format (bfseries, itshape, etc.)
  {}                       % label  (thesection, thesubsection, etc.)
  {0em}                    % separation between label and body
  {#1}        % before the body
  [.]                       % after the body

  \let\lil\lstinputlisting
  \usepackage{afterpage}
  \hypersetup{colorlinks, allcolors=., urlcolor=blue}

  \usepackage{etoolbox}% http://ctan.org/pkg/etoolbox
  \makeatletter
  \patchcmd{\lst@GLI@}% <command>
    {\def\lst@firstline{#1\relax}}% <search>
    {\def\lst@firstline{#1\relax}\def\lst@firstnumber{#1\relax}}% <replace>
    {\typeout{listings firstnumber=firstline}}% <success>
    {\typeout{listings firstnumber not set}}% <failure>
  \makeatother

  \renewcommand\lstlistlistingname{List of Listings}
  \usepackage{chngcntr}
  \counterwithin{figure}{section}
  \counterwithin{table}{section}

  \usepackage{booktabs}
  \usepackage{longtable}
  \usepackage{array}
  \usepackage{multirow}
  \usepackage{wrapfig}
  \usepackage{float}
  \usepackage{colortbl}
  \usepackage{pdflscape}
  \usepackage{tabu}
  \usepackage{threeparttable}
  \usepackage{threeparttablex}
  \usepackage[normalem]{ulem}
  \usepackage{makecell}
  \pagestyle{fancy}
  ```
output:
  pdf_document:
    number_sections: true
    fig_crop: true
    fig_caption: true
    keep_tex: false
---
```{r,include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(kableExtra.latex.load_packages = FALSE)
library(kableExtra)
```
\thispagestyle{empty}
\counterwithin{lstlisting}{section}
\newpage
\pagenumbering{roman}
\thispagestyle{plain}
\tableofcontents
\newpage
\thispagestyle{plain}
\listoffigures
\listoftables
\lstlistoflistings
\newpage
\pagenumbering{arabic}
\setcounter{page}{1}

# Introduction

## The Buddhist Parable of the Mustard Seed
- cosmological claim
  - souls "flow" for a hundred thousand ages
- Skinny Gotami
  - has a son so people respect her
  - son dies
  - sorrow to the point of madness
- desperately carries the body around begging for medicine
  - finally sent to the Buddha
- the Buddha tells her
  - find a family unaffected by death and get a mustard seed from them
  - Gotami eventually realizes everyone is affected by death
  - she finds enlightenment

## Characteristics of Religious Ritual
- rituals
  - invariant
  - understanding of an act
- religious rituals
  - performed
    - real audience
      - sermon
      - funeral
      - wedding
    - imagined audience
      - prayer
  - formalistic
    - governed by rules
  - traditionalistic
    - appears to not change
    - any change occurs very slowly
- ritualization posits the existence of an **authoritative reality**
  - dictates the immediate situation during a ritual

### Effects of Ritual
- forms/reinforces communities
- connects
  - actions
  - authoritative reality
- explanation of social realities
  - reverence of elders in Chinese religion
    - at a funeral, the youngest descendant pretends they are the deceased
    - they are shown the same amount of respect as the deceased elder
    - a preview of what's to come

\newpage

## Van Gennep's Rite of Passage
- always a **change in social status**
- three stages
  1. separation
     - taken out of typical social circumstances
  1. liminality
     - between two states
     - e.g., both a man and a boy at once
     - limited individuality
  1. reintegration
     - back to normal
     - often with a special sign of having completed ritual
       - wedding bands
       - etc.

## Different Definitions of Death

### Traditional
- cessation of breath

### Veatch's Four Possibilities
1. irreversible loss of **flow of vital fluids**
1. irreversible loss of **soul**
1. irreversible loss of **brain function** (See [Harvard Definition](#Harvard))
1. irreversible loss of **capacity for social function** due to **loss of brain function**

### Chinese
- Iron Crutch Li
  - visiting the gods
- perhaps the dead are just visiting the gods?
  - leave them out for a few days to make sure

### Harvard {#Harvard}
- irreversible loss of **brain function**
  - even if life mechanically supported
  - brain death
- this is the contemporary definition

## Issues With Comparative Study of Religion
- important to stay **objective**
  - subjective opinions lead to loss of focus
- keep things
  - phenomenological (See \autoref{glossary})
  - epoch&eacute; (See \autoref{glossary})
- focus on **comparison and contrast**
  - how does a given religion handle death?

### E.B. Tyler
- "primitive cultures"
  - somewhat pejorative
  - implies inferiority

### Essentialism
- assumption that there is **one initial religion**
  - all others would branch from it
- **zero evidence for this**

```{r child='./readings/week1.rmd'}
```

```{r child='./readings/week1-video.rmd'}
```





# Indigenous Religions

## The Role of Affect in Religion

## Basilov's Characteristics of Indigenous Religions

## Souls and Spirits in Indigenous Religions

## Wari' Funeral Practice

## The Melpa Understanding of Souls

```{r child='./readings/week2.rmd'}
```





# Death in the Ancient World

## The Egyptian View of the Soul

## Three Textual Primary Sources of Understanding the Egyptian Views on Death

## Death in "The Epic of Gilgamesh"

## Babylonian Funeral Practice

## Zoroastrian Funerals

## The Divisions of the Underworld in Greek Thought

## Early Human Perspectives on Death

## What Constitutes a Burial

```{r child='./readings/week3.rmd'}
```

\newpage
\appendix
# Definitions

```{r}
data = read.csv(file='./csv/definitions.csv', check.names=FALSE, stringsAsFactors=FALSE,
encoding='UTF-8')

knitr::kable(
  data, caption = '\\label{glossary}A glossary of important definitions.', longtable=TRUE, align=c("r","l")
) %>%
group_rows("Week 1: Introduction",1,10, indent=FALSE) %>%
group_rows("Week 2: Indigenous Religions",11,16, indent=FALSE) %>%
group_rows("Week 3: Death in the Ancient World",17,26) %>%
kable_styling(latex_options = c('striped'), stripe_color="#e6f0ff") %>%
column_spec(2,width='32em') %>%
column_spec(1,width='11em') %>%
row_spec(0,bold=TRUE)
```

\newpage

# Important Dates

```{r}
data = read.csv(file='./csv/dates.csv', check.names=FALSE, stringsAsFactors=FALSE,
encoding='UTF-8')

knitr::kable(
  data, caption = '\\label{dates}Important dates.', longtable=TRUE, align=c("r","r","l","l")
) %>%
kable_styling(latex_options = c('striped'), stripe_color="#e6f0ff") %>%
column_spec(3,width='30em') %>%
row_spec(0,bold=TRUE)
```
